What We Eval in the Shadows

### 2021-09-03 paper/main.tex
Fix typos and grammar in Section 3

\item \emph{CRAN.} Packages hosted on the Comprehensive R Archive Network
  ({\small \url{cran.r-project.org}}, aka CRAN) have three sources of runnable:
  unit \emph{tests} for individual functions, code \emph{examples} embedded in
  the documentation, and use-cases called \emph{vignettes}. All runnable
  snippets can be extracted into independent files. We download \CranPackages
  packages and extract \CranRunnableScripts scripts with \CranRunnableCode
  lines of code (\CranRunnableCodeExamplesRnd in examples,
  \CranRunnableCodeVignettesRnd in vignettes and \CranRunnableCodeTestsRnd in
  tests). CRAN packages contain 20 years of contribution from
  thousands of authors, often experienced R developers.

\item \emph{Kaggle.} Kaggle ({\small \url{kaggle.com}}) is an online platform
  that allows users to submit and compete to solve problems. We download
  \KaggleUnique unique scripts and notebooks with their input files
  (\KaggleDuplicates duplicates were removed using SHA-1 hashes); 665K lines of
  R code making only \KaggleWithEvals calls \eval. Authors of these scripts
  have highly variable levels of expertise.

@@

We implemented an automated pipeline that acquires packages, extracts scripts,
executes them, traces their behavior, and summarizes observations.
Figure~\ref{fig:pipeline} shows the main steps along with their running time,
data size, and number of elements manipulated. The pipeline steps are:

@@

\begin{compactenum}[(1)]
\item \emph{Download.} Packages are downloaded from CRAN, for Kaggle a web
  crawler retrieves code, and a command-line tool gets data. Installation is
  complicated by native dependencies which sometimes have to be resolved
  manually.
\item \emph{Extract.} The \genthat tool extracts all runnable code snippets and
  turns each of these into a self-standing program~\cite{issta18}, \k{knitr}
  extracts code from notebooks. Each extracted script is instrumented with calls
  to our dynamic analyzer in a way that avoid recording calls to \eval
  originating from our execution harness.
\item \emph{Trace.} Script run with a modified R interpreter to capture calls
  to \eval in CRAN packages. Each script is run in its own process with the GNU
  R compiler turned off to avoid recording its execution.
\item \emph{Analyze.} Analysis outputs are merged, cleaned, and summarized.
  RMarkdown notebooks process the summarized data to generate all graphs (PDF
  from \k{ggplot2}) and numbers (exported as \LaTeX macros) appearing in the
  paper.
\end{compactenum}

@@

The tracer that performs dynamic analysis of R scripts is built on top of
\rdyntrace, an extended R 4.0.2 virtual machine that exposes callbacks for
various runtime events~\cite{oopsla19b}. The tracer registers callbacks to all
variants of \eval and a few additional functions to locate the origin of \eval
arguments. It captures dynamic code loading, as well as variable definition and
assignment, allowing us to record side effects that happen in environments
while evaluating code in \eval. A challenge was the lack of source code
references for expression that are not within block surrounded by braces. This
is unfortunately not easily fixed. We extend our dynamic analysis tool to
attach synthetic source code references to all \eval call sites, but the
approach fails in some edge cases. For performance reasons, the tracer is an R
package written in C++ (3.2K LOC) and R (1.3K LOC). While in theory the
implementation should be straightforward, not so in practice. Lazy evaluation
requires delaying the processing of arguments until (and if) they are forced.
Accounting for side-effects performed in an \eval is complicated by the fact
that R is implemented in a mixture of R and C, and that the language
implementation can (and does) call \eval. Lastly, large code bases exercise
many edge cases of the highly underspecified R behavior.

\paragraph{Limitations.} Even with the extension described above, there are
\PkgUndefinedRnd \eval calls without source information (\PkgUndefinedRatio of
all \eval calls). This happen when \eval is an argument to a higher-order
function or when it is called from native code. Another limitation is that we
do not record calls to the native \eval. Neither of these limitations should
invalidate our conclusions.

>>>>

\item \emph{CRAN.} Packages hosted on the Comprehensive R Archive Network
  ({\small \url{cran.r-project.org}}, aka CRAN) have three sources of runnable code:
  unit \emph{tests} for individual functions, code \emph{examples} embedded in
  the documentation, and use-cases called \emph{vignettes}. All runnable
  snippets can be extracted into independent files. We download \CranPackages
  packages and extract \CranRunnableScripts scripts with \CranRunnableCode
  lines of code (\CranRunnableCodeExamplesRnd in examples,
  \CranRunnableCodeVignettesRnd in vignettes and \CranRunnableCodeTestsRnd in
  tests). CRAN packages contain 20 years of contribution from
  thousands of authors, often experienced R developers.

\item \emph{Kaggle.} Kaggle ({\small \url{kaggle.com}}) is an online platform
  that allows users to submit and compete to solve problems. We download
  \KaggleUnique unique scripts and notebooks with their input files
  (\KaggleDuplicates duplicates were removed using SHA-1 hashes); 665K lines of
  R code making only \KaggleWithEvals calls \eval. The authors of these scripts
  have highly variable levels of expertise.

@@

We implemented an automated pipeline that acquires packages, extracts scripts,
executes them, traces their behavior, and summarizes observations.
Figure~\ref{fig:pipeline} shows the main steps along with their running time,
data size, and the number of elements manipulated. The pipeline steps are:

@@

\begin{compactenum}[(1)]
\item \emph{Download.} Packages are downloaded from CRAN, for Kaggle a web
  crawler retrieves code, and a command-line tool gets data. Installation is
  complicated by native dependencies which sometimes have to be resolved
  manually.
\item \emph{Extract.} The \genthat tool extracts all runnable code snippets and
  turns each of these into a self-standing program~\cite{issta18}, \k{knitr}
  extracts code from notebooks. Each extracted script is instrumented with calls
  to our dynamic analyzer in a way that avoids recording calls to \eval
  originating from our execution harness.
\item \emph{Trace.} Scripts are run with a modified R interpreter to capture
  calls to \eval in CRAN packages. Each script is run in its own process with
  the GNU R compiler turned off to avoid recording its execution.
\item \emph{Analyze.} Analysis outputs are merged, cleaned, and summarized.
  RMarkdown~\cite{rmarkdown} notebooks process the summarized data to generate all graphs (using
  \k{ggplot2}~\cite{ggplot}) and numbers (exported as \LaTeX\xspace macros)
  appearing in the paper.
\end{compactenum}

@@

The tracer that performs dynamic analysis of R scripts is built on top of
\rdyntrace, an extended R 4.0.2 virtual machine that exposes callbacks for
various runtime events~\cite{oopsla19b}. The tracer registers callbacks to all
variants of \eval and a few additional functions to locate the origin of \eval
arguments. It captures dynamic code loading, as well as variable definition and
assignment, allowing us to record side effects that happen in environments while
evaluating code in \eval. A challenge was the lack of source code references for
expressions that are not within a block surrounded by braces. This is
unfortunately not easily fixed. We extend our dynamic analysis tool to attach
synthetic source code references to all \eval call sites, but the approach fails
in some edge cases. For performance reasons, the tracer is an R package written
in C++ (3.2K LOC) and R (1.3K LOC). While in theory, the implementation should
be straightforward, it is not so in practice. Lazy evaluation requires delaying
the processing of arguments until (and if) they are forced. Accounting for
side-effects performed in an \eval is complicated by the fact that R is
implemented in a mixture of R and C, and that the language implementation can
(and does) call \eval. Lastly, large codebases exercise many edge cases of the
highly underspecified R behavior.

\paragraph{Limitations.} Even with the extension described above, there are
\PkgUndefinedRnd \eval calls without source information (\PkgUndefinedRatio of
all \eval calls). This happens when \eval is an argument to a higher-order
function or when it is called from native code. Another limitation is that we do
not record calls to the native \eval. Neither of these limitations should
invalidate our conclusions.
